Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-24 VMCAI-revised-version/final.tex
effectivity

\alain{remercier les reviewers !} We thank the reviewers of the VMCAI 2024 conference for their numerous and relevant comments and improvement suggestions.

@@

\begin{proof}
\alain{preuve pas claire, à réécrire}
% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
%which can be simulated by WSTS.
First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
	This is due to the undecidability of $2$-counter Minsky machine termination.
From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$.
Based on this construction, the problem of deciding whether the downward-closed set 
$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable in the $3$-counter Minsky machine $M'$
% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable~\cite{Min61}. 
Note that by chosing $s_0 \in Q_{M'} \times \mathop{\uparrow} (0,0,1)$,
deciding whether the downward-closed set 
$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable
from every element of $\post^*(s_0) \cup Q_{M'} \times \mathop{\uparrow} (0,0,1)$ 
is again equivalent to asking whether $M$ terminate on every input.
Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.

>>>>

\alain{remercier les reviewers !} We express our thanks to the reviewers of the VMCAI 2024 Conference for their numerous and relevant comments and improvement suggestions.

@@

\begin{proof}
\alain{preuve pas claire, à réécrire}
% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
%which can be simulated by WSTS.
First, let us show that {\sc Resilience} is undecidable for $3$-counter machines.
%	This is due to the undecidability of termination for $2$-counter Minsky machines.
From an $2$-counter machine $M=(Q, q_f,...)$ where $Q$ is the set of control-states and $q_f \in Q$ is the final control-state, one constructs an associated $3$-counter machine $M'=(Q', q'_f,...)$ with a final control-state $q_f' \in Q'$
such that $M$ terminates on all initial states iff $q_f'(0,0,0)$ is reachable (in $M'$) from any state in $Q' \times \mathop{\uparrow} (0,0,1)$. 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches $q_f$, then $M'$ decreases its first two counters until they reach $0$, then, and only then, finally $M'$ decreases the third counter until it reaches $0$.
%
Remark that $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$.
%
Based on this construction, the downward-closed set 
$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable (in $M'$)
% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ iff $M$ terminates on all initial states; this last problem is undecidable~\cite{Min61}. 
Note that by chosing $s_0 \in Q' \times \mathop{\uparrow} (0,0,1)$,
deciding whether the downward-closed set 
$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable
from every state in $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1)$ 
is again equivalent to asking whether $M$ terminates on every initial state.
Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
